---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%",
  dev = "png", dpi = 300,
  fig.height = 3.5, 
  cache = FALSE
)
```

# **ggvfields**

## Overview

**ggvfields** provides tools for visualizing vector fields, complex numbers, and soon dual numbers and more. 

```{r}
library("ggvfields")
```

## Usage

### `geom_vector_field()`{.R}

The `geom_vector_field()`{.R} function generates a vector field plot layer using a user-defined function to compute the vector components. This function abstracts away the mathematical computations required to generate the vector field, so the user does not need to manually calculate and input the vector components into `geom_segment()`{.R}. It simplifies the process, making it easier to create vector field visualizations without dealing with the underlying math.

```{r}
# custom function
f <- function(v) {
  x <- v[1]; y <- v[2]
  c(-y, x) # = f(x,y)
}

# create a ggplot with the vector field layer
ggplot() +
  geom_vector_field(fun = f, xlim = c(-10, 10), ylim = c(-10, 10)) +
  coord_fixed()
```

By default, the magnitude (or length/norm) of the vector lengths is mapped to the color aesthetic. 

The norm of a vector $\mathbf{w} = (u, v)$ is given by:

$|\mathbf{w}| = \sqrt{u^2 + v^2}$

This function also offers several other aesthetic mappings which allow you to map several characteristics of the vector field to different aesthetic mappings.

#### Divergence

Divergence is an operation on a vector field that tells us how the field behaves toward or away from a point. Locally, the divergence of a vector field $\mathbf{F}$ in $\mathbb{R}^2$ at a particular point is a measure of the "outflowing-ness" of the vector field at that particular point. The divergence of a vector field results in a scalar function.

If $\mathbf{F} = \langle \mathbf{F}_x(x,y), \mathbf{F}_y(x,y) \rangle$ is a vector field in $\mathbb{R}^2$, then the divergence of $\mathbf{F}$ is defined by:

$$
\text{div} \, \mathbf{F} = \frac{\partial \mathbf{F}_x}{\partial x} + \frac{\partial \mathbf{F}_y}{\partial y}
$$

To visualize the divergence of the vector field:

```{r}
ggplot() +
  geom_vector_field(
    aes(color = after_stat(divergence)), 
    fun = f, xlim = c(-10, 10), ylim = c(-10, 10)
  ) +
  coord_fixed()
```

#### Curl

Curl is an operation on a vector field that measures the rotation or swirling strength at a point. In two dimensions, the curl of a vector field $\mathbf{F}$ in $\mathbb{R}^2$ is a scalar value that indicates how much the vector field tends to rotate around that point.

If $\mathbf{F} = \langle \mathbf{F}_x(x,y), \mathbf{F}_y(x,y) \rangle$ is a vector field in $\mathbb{R}^2$, then the curl of $\mathbf{F}$ is defined by:

$$
\text{curl} \, \mathbf{F} = \frac{\partial \mathbf{F}_y}{\partial x} - \frac{\partial \mathbf{F}_x}{\partial y}
$$

To visualize the magnitude of the curl:

```{r}
ggplot() +
  geom_vector_field(
    aes(color = after_stat(curl)), 
    fun = f, xlim = c(-10, 10), ylim = c(-10, 10)
  ) +
  coord_fixed()
```

#### Laplace Operator

The Laplace operator (or Laplacian) measures the rate at which the average value of a function around a point differs from the value at that point. It is commonly used to understand how a scalar field spreads out or compresses in space.

If $\mathbf{F} = \langle \mathbf{F}_x(x,y), \mathbf{F}_y(x,y) \rangle$ is a vector field in $\mathbb{R}^2$, then the Laplacian of $\mathbf{F}$ is defined by taking the sum of the second partial derivatives of each component of the vector field.

The Laplacian of the vector field $\mathbf{F}$ is thus given by:

$$
\Delta \mathbf{F} = \frac{\partial^2 \mathbf{F}_x}{\partial x^2} + \frac{\partial^2 \mathbf{F}_x}{\partial y^2} + \frac{\partial^2 \mathbf{F}_y}{\partial x^2} + \frac{\partial^2 \mathbf{F}_y}{\partial y^2}
$$

This results in a scalar value that describes how the vector field spreads out or compresses at different points in the field.

To visualize the Laplace operator of the vector field:

```{r}
ggplot() +
  geom_vector_field(
    aes(color = after_stat(laplacian)), 
    fun = f, xlim = c(-10, 10), ylim = c(-10, 10)
  ) +
  coord_fixed()
```

#### Directional Derivative

The directional derivative of a vector field measures the rate of change of the field in a specified direction. It provides insight into how the field varies as you move in a particular direction.

If $\mathbf{F} = \langle \mathbf{F}_x(x,y), \mathbf{F}_y(x,y) \rangle$ is a vector field in $\mathbb{R}^2$, and $\mathbf{v} = \langle v_x, v_y \rangle$ is a direction vector, then the directional derivative of $\mathbf{F}$ in the direction of $\mathbf{v}$ is given by:

$$
D_{\mathbf{v}} \mathbf{F} = \frac{\partial \mathbf{F}_x}{\partial x} v_x + \frac{\partial \mathbf{F}_x}{\partial y} v_y + \frac{\partial \mathbf{F}_y}{\partial x} v_x + \frac{\partial \mathbf{F}_y}{\partial y} v_y
$$

For a vector field, this represents the rate of change of each component of the field in the direction specified by $\mathbf{v}$.

To visualize the Laplace operator of the vector field:


```{r}
x1 <- 5
y1 <- 6

ggplot() +
  geom_vector_field(
    aes(color = after_stat(directional_derivative)),
    fun = f, xlim = c(-10, 10), ylim = c(-10, 10),
    u = x1, v = y1) +
  geom_point(aes(x = x1, y = y1)) +
  scale_color_viridis_c() +
  coord_fixed()
```

### `geom_complex_function()`{.R}

The `geom_complex_function()`{.R} function generates a vector field plot layer using a user-defined function to compute the vector components. This function abstracts away the mathematical computations required to generate the vector field, so the user does not need to manually calculate and input the vector components into `geom_segment()`{.R}. It simplifies the process, making it easier to create vector field visualizations without dealing with the underlying math.

```{r}
# custom function
f <- function(z) (z^2 + 1) / (z^2 - 1)

# create a ggplot with the complex vector field layer
ggplot() +
  geom_complex_function(fun = f, relim = c(-2, 2), imlim = c(-2, 2), n = 100) +
  coord_fixed()
```


## License

This package is licensed under the MIT License.


## Contact

For any questions or issues, please [open an issue](https://github.com/dusty-turner/ggvfields/issues/new) on GitHub or contact the maintainer.


## Installation


You can install the development version from GitHub:

```{r, eval=FALSE}
devtools::install_github("dusty-turner/ggvfields")
```


## Related projects

**ggvfields** is similar to [**ggquiver**](http://pkg.mitchelloharawild.com/ggquiver/).
