---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%",
  dev = "png", dpi = 300,
  fig.height = 3.5, 
  cache = FALSE
)
```

# **ggvfields**

## Overview

**ggvfields** provides tools for visualizing vector fields, stream plots, and soon complex numbers and more.

```{r eval = FALSE}
remotes::install_github("dusty-turner/ggvfields")
```


```{r}
library("ggvfields")
options(ggplot2.continuous.colour="viridis")
```

## Usage

### `geom_vector_field()`{.R}

The `geom_vector_field()`{.R} function generates a vector field plot layer using a user-defined function to compute the vector components. This function abstracts away the mathematical computations required to generate the vector field, so the user does not need to manually calculate and input the vector components into `geom_segment()`{.R}. It simplifies the process, making it easier to create vector field visualizations without dealing with the underlying math.

```{r}
f <- function(v) {
  x <- v[1]; y <- v[2]
  c(-y, x) # = f(x,y)
}

ggplot() +
  geom_vector_field(fun = f, xlim = c(-10, 10), ylim = c(-10, 10)) +
  coord_fixed() 
```

This function allows the user to map several characteristics of the vector field to different aesthetic mappings.

#### Norm

The norm of a vector $\mathbf{w} = (u, v)$ is given by:

$|\mathbf{w}| = \sqrt{u^2 + v^2}$

We can visualize the norm by mapping the value of the norm to the color aesthetic.

```{r}
ggplot() +
  geom_vector_field(
    aes(color = after_stat(norm)),
    fun = f, xlim = c(-10, 10), ylim = c(-10, 10)) +
  coord_fixed()
```


#### Divergence

Divergence is an operation on a vector field that tells us how the field behaves toward or away from a point. Locally, the divergence of a vector field $\mathbf{F}$ in $\mathbb{R}^2$ at a particular point is a measure of the "outflowing-ness" of the vector field at that particular point. The divergence of a vector field results in a scalar function.

If $\mathbf{F} = \langle \mathbf{F}_x(x,y), \mathbf{F}_y(x,y) \rangle$ is a vector field in $\mathbb{R}^2$, then the divergence of $\mathbf{F}$ is defined by:

$$
\text{div} \, \mathbf{F} = \frac{\partial \mathbf{F}_x}{\partial x} + \frac{\partial \mathbf{F}_y}{\partial y}
$$

To visualize the divergence of the vector field:

```{r}
ggplot() +
  geom_vector_field(
    aes(color = after_stat(divergence)), 
    fun = f, xlim = c(-10, 10), ylim = c(-10, 10)
  ) +
  coord_fixed()
```

#### Curl

Curl is an operation on a vector field that measures the rotation or swirling strength at a point. In two dimensions, the curl of a vector field $\mathbf{F}$ in $\mathbb{R}^2$ is a scalar value that indicates how much the vector field tends to rotate around that point.

If $\mathbf{F} = \langle \mathbf{F}_x(x,y), \mathbf{F}_y(x,y) \rangle$ is a vector field in $\mathbb{R}^2$, then the curl of $\mathbf{F}$ is defined by:

$$
\text{curl} \, \mathbf{F} = \frac{\partial \mathbf{F}_y}{\partial x} - \frac{\partial \mathbf{F}_x}{\partial y}
$$

To visualize the the curl:

```{r}
ggplot() +
  geom_vector_field(
    aes(color = after_stat(curl)), 
    fun = f, xlim = c(-10, 10), ylim = c(-10, 10)
  ) +
  coord_fixed()
```

#### Aesthetics In Development

Another aesthetic in development available for mapping is length.  The length aesthetic maps the value of the computed measure to the length of the vector.

```{r}
ggplot() +
  geom_vector_field(
    aes(length = after_stat(norm), color = after_stat(norm)),
    fun = f, xlim = c(-10, 10), ylim = c(-10, 10)) +
  coord_fixed() +
  theme(legend.box = "horizontal")
```

### `geom_vector()`{.R}

The `geom_vector()`{.R} function generates a plot layer that visualizes vectors as line segments with optional arrowheads. It is designed to help you visualize directional data, such as wind directions, gradients, or flow fields. The vectors are defined by their start (`x`, `y`) and end (`xend`, `yend`) coordinates, or alternatively by their angular (`angle`) and distance (`distance`) components.

#### Cartesian

```{r}
set.seed(1234)
n <- 10
wind_data_polar <- data.frame(
  lon = rnorm(n), 
  lat = rnorm(n), 
  wind_dir = runif(n, -pi, pi),
  wind_spd = rchisq(n, df = 2)
)

wind_data_cartesian <- within(wind_data_polar, {
  wind_lon_comp <- wind_spd * cos(wind_dir)
  wind_lat_comp <- wind_spd * sin(wind_dir)
  xend <- lon + wind_lon_comp
  yend <- lat + wind_lat_comp
})

wind_data_cartesian |> 
  ggplot() +
  geom_vector(aes(x = lon, y = lat, xend = xend, yend = yend)) +
  labs(title = "Wind Vectors (Cartesian Input)",
       x = "Longitude", y = "Latitude")
```

#### Polar

```{r}

wind_data_polar |> 
  ggplot() +
  geom_vector(aes(x = lon, y = lat, angle = wind_dir * 180 / pi, distance = wind_spd)) +
  labs(title = "Wind Vectors (Polar Input)",
       x = "Longitude", y = "Latitude")

```



### `geom_streamplot()`{.R}

The `geom_streamplot()`{.R} function generates a stream plot layer of a user-defined vector field function. The lines in the plot represent the flow of data points through the vector field.

```{r}
f <- function(v) {
  x <- v[1]
  y <- v[2]
  c(-1 - x^2 + y, 1 + x - y^2)
}

ggplot() +
  geom_streamplot(
    fun = f, xlim = c(-3, 3), ylim = c(-3, 3), 
    ) +
  coord_fixed() +
  theme_minimal()
```

The `chop` parameter (defaulted to TRUE) allows you to chop the trajectories into segments. This can be useful for better visualization of the streamlines when they are long and complex. 

It may be useful to not break up the streamlines.

```{r}
ggplot() +
  geom_streamplot(
    fun = f, xlim = c(-3, 3), ylim = c(-3, 3), 
    chop = FALSE
    ) +
  coord_fixed() +
  theme_minimal()
```

It may also be useful to break up the streamlines into more segments. The `scale_stream` parameter (defaults to 1) adjusts the segmentation of streamlines by specifying the proportion of the streamline length used to divide it into smaller segments.

```{r}
ggplot() +
  geom_streamplot(
    fun = f, xlim = c(-3, 3), ylim = c(-3, 3), 
    chop = TRUE, scale_stream = .9,
    ) +
  coord_fixed() +
  theme_minimal()
```

### Map Calculus Measures to Aesthetics

```{r}
ggplot() +
  geom_streamplot(
    fun = f, xlim = c(-3, 3), ylim = c(-3, 3), 
    aes(color = after_stat(log(divergence + 10)))
    ) +
  coord_fixed() +
  theme_minimal() 
```



### Animate `geom_streamplot()`{.R}

```{r}
#| cache: true
# Create stream plot with rownum aesthetic

p <- ggplot() +
 geom_streamplot(
   aes(rownum = after_stat(rownum)), 
   fun = f, xlim = c(-3, 3), ylim = c(-3, 3)
 ) +
 coord_fixed() +
 theme_bw()

# Create an animation transition plot
anim <- animation_transition(plot = p) +   
 gganimate::transition_reveal(rownum) +
 gganimate::ease_aes('linear')

# Animate the plot
gganimate::animate(anim, nframes = 25, fps = 5, end_pause = 0, renderer = gganimate::gifski_renderer())
```


The `mask_shape_type` parameter allows you to specify the mask shape used for streamline generation which influences how the streamlines are placed and how closely they can approach each other. The default mask shape is `"square"`, but you can also use `"diamond"`, `"inset_square"`, or `"circle"`.  During streamline generation, when a streamline enters the specified shape, no other streamlines will enter that region. 

- **Square Mask (default)**: Streamlines are restricted to a grid where each cell is a square. This generally results in evenly spaced streamlines.
- **Diamond Mask**: Streamlines are restricted to a square grid with diamonds inset within each square.  This can create a more dense pattern which can have better visualizations for some functions. - **Inset Square Mask**: Streamlines are restricted to a grid with smaller squares inset within larger squares. This can create a denser and more detailed pattern of streamlines.
- **Diamond Mask**: Streamlines are restricted to a square grid with diamonds inset within each square.  This can create a more dense pattern which can have better visualizations for some functions. - **Circle Mask**: Streamlines are restricted to a grid with inset circles inside the square grid.


```{r}
ggplot() +
  geom_streamplot(aes(group = after_stat(id)),
                  fun = f, xlim = c(-3, 3), ylim = c(-3, 3), max_length = 10000,
                  max_steps = 10000, ds = .05, min_dist = .25, 
                  mask_shape_type = "diamond") +
  coord_fixed() +
  theme_minimal()
```

### `geom_complex_function()`{.R}

The `geom_complex_function()`{.R} function generates a vector field plot layer using a user-defined function to compute the vector components. This function abstracts away the mathematical computations required to generate the vector field, so the user does not need to manually calculate and input the vector components into `geom_segment()`{.R}. It simplifies the process, making it easier to create vector field visualizations without dealing with the underlying math.

```{r}
f <- function(z) (z^2 + 1) / (z^2 - 1)

ggplot() +
  geom_complex_function(fun = f, relim = c(-2, 2), imlim = c(-2, 2), n = 100) +
  labs(x = "Real", y = "Imaginary") +
  coord_fixed() +
  theme(legend.box = "horizontal")
```

We can enhance this plot with a little help from biscale.

Using biscale, we can apply a bivariate color scale to the plot, which allows us to represent two variables—angle (direction) and magnitude (intensity)—simultaneously. This makes it easier to visualize how these properties change across the field.

```{r}
library(biscale)
library(cowplot)

plot <-
  ggplot() +
  geom_complex_function(aes(fill = after_stat(bi_class)),
    fun = f, relim = c(-2, 2), imlim = c(-2, 2), n = 100
    ) +
    bi_scale_fill(pal = "DkBlue") +
    labs(
      title = "",
      x = "Real (Re)",
      y = "Imaginary (Im)"
    ) +
    bi_theme(base_size = 16) +
    theme(legend.position = "none") +
  coord_fixed() 
  
legend <- bi_legend(pal = "DkBlue",
                    xlab = "Angle",
                    ylab = "Magnitude",
                    size = 6)

ggdraw() +
  draw_plot(plot, 0, 0, .8, 1) +  
  draw_plot(legend, x = .55, y = .6, width = .3, height = 0.3)
```


## License

This package is licensed under the MIT License.

## Contact

For any questions or issues, please [open an issue](https://github.com/dusty-turner/ggvfields/issues/new) on GitHub or contact the maintainer.

## Installation

You can install the development version from GitHub:

```{r, eval=FALSE}
devtools::install_github("dusty-turner/ggvfields")
```

## Related projects

For creating vector fields, the [**ggquiver**](http://pkg.mitchelloharawild.com/ggquiver/) package provides quiver plots to visualize vector fields, while the [**ggarchery**](https://github.com/mdhall272/ggarchery) package can handling of segments with arrowheads.
